Network based multifactorial modelling microRNA-target interaction.

### 2019-10-08 NAR_form_article.Rmd
delete todos

## Breast cancer patient dataset

We have applied our model in a real dataset for which experimental measurements of various factors were available. Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network https://www.cancer.gov/tcga. High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015] (Detailed codes are available at Section 6 in Supplementary Materials and Methods). We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015] (Section 6 in Supplementary Materials and Methods and Table S4 in Supplementary Tables). SERPINE2, one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network. Relevant steps and codes are available at Supplementary Materials and Methods.

[#TODO]: <> "about formulas, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?"

@@

[#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur"

@@

# ACKNOWLEDGEMENTS

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text.

>>>>

## Breast cancer patient dataset
We have applied our model in a real dataset for which experimental measurements of various factors were available. Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network https://www.cancer.gov/tcga. High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015] (Detailed codes are available at Section 6 in Supplementary Materials and Methods). We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015] (Section 6 in Supplementary Materials and Methods and Table S4 in Supplementary Tables). SERPINE2, one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network. Relevant steps and codes are available at Supplementary Materials and Methods.

@@

@@

### 2019-10-08 NAR_form_article.tex
eps convertion of pics
delete todos

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=15cm,height=8cm]{Fig3.jpg}
\end{center}
\caption{Target regulations with interaction parameters. 
     (\textbf{A}) In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. 
     (\textbf{B}) The changes the repression activities after increasing of G2 expression. 
     (\textbf{C}) Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). 
     (\textbf{D}) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2)}
\label{fig:fig3}
\end{figure*}

@@

Earlier studies reported that miRNA regulatory interactions are affected
by different parameters. For example, Xu et al.~have investigated the
importance of seed pairing type between miRNAs and their targets and
target site location by using proteomics dataset
\citep{xu_characterization_2014}. They have proposed that the features
of binding between miRNA and target can be critical for miRNA
efficiency. In addition, binding energy between miRNA and targets is a
significant determinant for miRNA efficiency and it has been reported
that strength of miRNA:target interactions is depended on binding energy
of complexes \citep{breda_quantifying_2015}. Similarly, another study
revealed that affinity is correlated with seed pairing of miRNA:target
pairs and suggested that affinity is correlated with number of canonical
seed base pairing \citep{bosson_endogenous_2014}. Therefore, we have
integrated aforementioned interaction parameters that could be useful
for predicting miRNA repression activity more accurately. The sample
dataset used in Figure \ref{fig:fig1} is recalculated with additional
factors Table \ref{tab:one} in effect with same trigger (Detailed codes
can be found at Section 3 in Supplementary Materials and Methods), two
fold increase in Gene2 (Figure \ref{fig:fig3}B, and Figure S6). When the
factors were taken into account in the system, miRNA efficiencies varied
as shown in Figure \ref{fig:fig3}A (Figure S5). Although the
miRNA:target expression ratios in steady-state were same in comparison
with the sample dataset without factors, efficiency of binding and
repression have changed. On the other hand, changes in expression levels
of G4 (Detailed codes are available at Section 4 in Supplementary
Materials and Methods; and Figure S8) and G1 differ although they have
same initial conditions. This is due to fact that G4 is targeted by two
miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and changed interaction
factors.

@@

We have analysed the realistic network in context of perturbation
efficiencies of nodes. This process were run with 3 fold upregulation
and 10 iteration in conditions where elements that have less than five
percent change are not taken into account (Codes are found at Section
6.5 in Supplementary Materials and methods). With this approach,
SERPINE2 has been found as one of the most effective gene nodes in
network. We have observed that gaining steady-state in realistic network
takes a lot of time because of complex interactions, as expected, and
SERPINE2 exhibit highest efficiency at iteration 31 with given
conditions (see Section 6.6 in Supplementary Materials and Methos and
Figure S9 in Supplementary Figures). We ran the model with two fold of
iteration which is observed as most effective. We have observed that the
results of simulation with SERPINE2 gene, a serine protease inhibitor,
one of the overexpressed genes of breast cancer
\citep{yang2018expression, candia2006protease} is not convenient with
tumor tissue expression values, when we compared the tumor gene
expression dataset retrieved from TCGA and simulation results obtained
via our approach (see Section 6.6 in Supplementary Materials and
methods). Simulation results were not expected to be consistent with
tumor tissue expression values because large number of abnormal
regulations were observed in these tissues. So, it might not said that a
single gene is responsible for entire regulations in cancer cells.
Besides, the all other factors such as up/down regulation of miRNAs or
other genes were ignored. Therefore, it would be more useful to test our
approach with an ideal dataset that includes entire informations of gene
expression values before and after the regulation of a known gene at
existing tissue, and miRNA expression values of this tissue at initial
conditions.

@@

\section{ACKNOWLEDGEMENTS}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text.

>>>>

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=15cm,height=8cm]{Fig3.eps}
\end{center}
\caption{Target regulations with interaction parameters. 
     (\textbf{A}) In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. 
     (\textbf{B}) The changes the repression activities after increasing of G2 expression. 
     (\textbf{C}) Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). 
     (\textbf{D}) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2)}
\label{fig:fig3}
\end{figure*}

@@

Earlier studies reported that miRNA regulatory interactions are affected
by different parameters. For example, Xu et al.~have investigated the
importance of seed pairing type between miRNAs and their targets and
target site location by using proteomics dataset
\citep{xu_characterization_2014}. They have proposed that the features
of binding between miRNA and target can be critical for miRNA
efficiency. In addition, binding energy between miRNA and targets is a
significant determinant for miRNA efficiency and it has been reported
that strength of miRNA:target interactions is depended on binding energy
of complexes \citep{breda_quantifying_2015}. Similarly, another study
revealed that affinity is correlated with seed pairing of miRNA:target
pairs and suggested that affinity is correlated with number of canonical
seed base pairing \citep{bosson_endogenous_2014}. Therefore, we have
integrated aforementioned interaction parameters that could be useful
for predicting miRNA repression activity more accurately. The sample
dataset used in Figure \ref{fig:fig1} is recalculated with additional
factors Table \ref{tab:one} in effect with same trigger (Detailed codes
can be found at Section 3 in Supplementary Materials and Methods), two
fold increase in Gene2 (Figure \ref{fig:fig3}B, and Figure S6). When the
factors were taken into account in the system, miRNA efficiencies varied
as shown in Figure \ref{fig:fig3}A (Figure S5). Although the
miRNA:target expression ratios in steady-state were same in comparison
with the sample dataset without factors, efficiency of binding and
repression have changed. On the other hand, changes in expression levels
of G4 (Detailed codes are available in Figure S8 and Section 4 in
Supplementary Materials and Methods) and G1 differ although they have
same initial conditions. This is due to fact that G4 is targeted by two
miRNAs, M1 and M2 (Figure \ref{fig:fig3}D) and changed interaction
factors.

@@

We have analysed the realistic network in context of perturbation
efficiencies of nodes. This process was run with 3 fold upregulation and
10 iteration in conditions where elements that have less than five
percent change are not taken into account (Codes are found at Section
6.5 in Supplementary Materials and Methods). With this approach,
SERPINE2 has been found as one of the most effective gene nodes in
network. We have observed that gaining steady-state in realistic network
takes a lot of time because of complex interactions, as expected, and
SERPINE2 exhibit highest efficiency at iteration 31 with given
conditions (see Section 6.6 in Supplementary Materials and Methos and
Figure S9 in Supplementary Figures). We ran the model with two fold of
iteration which is observed as most effective. We have observed that the
results of simulation with SERPINE2 gene, a serine protease inhibitor,
one of the overexpressed genes of breast cancer
\citep{yang2018expression, candia2006protease} is not convenient with
tumor tissue expression values, when we compared the tumor gene
expression dataset retrieved from TCGA and simulation results obtained
via our approach (see Section 6.6 in Supplementary Materials and
Methods). Simulation results were not expected to be consistent with
tumor tissue expression values because large number of abnormal
regulations were observed in these tissues. So, it might not said that a
single gene is responsible for entire regulations in cancer cells.
Besides, the all other factors such as up/down regulation of miRNAs or
other genes were ignored. Therefore, it would be more useful to test our
approach with an ideal dataset that includes entire informations of gene
expression values before and after the regulation of a known gene at
existing tissue, and miRNA expression values of this tissue at initial
conditions.

@@
